---
title: "Assignment 04"
author: "Wenjie Deng"
format: 
  html:
    embed-resources: true
execute:
  warning: false
  message: false
---
```{r}

#| message: false

#| warning: false

library(tidyverse)

```
## Data
The data used in this analysis come from the [CMS Major Eligibility Group Information for Medicaid and CHIP Beneficiaries by Month](https://data.medicaid.gov/dataset/ea9b7db3-db71-4663-b4e1-67e11d1d4fcc).
Enrollment counts reported as “DS” were treated as missing, and observations flagged as unusable were excluded from the analysis.
```{r}
medicaid <- read_csv(

  "/Users/wenjenny/Desktop/MajorEligibilityGroup-montly-102025.csv"

) |>

  mutate(

    CountEnrolled = parse_number(CountEnrolled),

    Month = as.Date(paste0(Month, "01"), format = "%Y%m%d")

  ) |>

  filter(

    !is.na(CountEnrolled),

    dunusable != "Unusable" | is.na(dunusable)

  )


medicaid |>
  count(`majoreligibility group`)
```

## Visualization 1: Enrollment by Eligibility Group
```{r}
latest_month <- max(medicaid$Month)

enrollment_group <- medicaid |>
  filter(
    Month == latest_month,
    `majoreligibility group` != "Unknown"
  ) |>
  group_by(`majoreligibility group`) |>
  summarize(
    enrollment = sum(CountEnrolled)
  )

enrollment_group |>
  ggplot(
    aes(
      x = reorder(`majoreligibility group`, enrollment),
      y = enrollment,
      fill = `majoreligibility group`
    )
  ) +
  geom_col(width = 0.7) +
  geom_text(
    aes(
      label = scales::label_number(
        scale = 1e-6,
        suffix = "M",
        accuracy = 0.1
      )(enrollment)
    ),
    hjust = -0.1,
    size = 3.5
  ) +
  coord_flip() +
  scale_y_continuous(
    labels = scales::label_number(
      scale = 1e-6,
      suffix = "M"
    ),
    expand = expansion(mult = c(0, 0.18))
  ) +
  labs(
    title = "Children Make Up the Largest Medicaid and CHIP Eligibility Group",
    subtitle = "Enrollment by major eligibility group across reporting jurisdictions, December 2022",
    x = NULL,
    y = "Enrollment",
    fill = "Eligibility group",
    caption = "Source: Centers for Medicare & Medicaid Services (CMS)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none"
  )
```

Children represented the largest Medicaid and CHIP eligibility group in December 2022, with approximately 43.1 million enrollees. The adult expansion group was the second largest, with about 25.5 million enrollees. Enrollment among COVID newly-eligible beneficiaries was substantially smaller than in the other major eligibility groups.
```
Children represented the largest Medicaid and CHIP eligibility group in December 2022, with approximately 43.1 million enrollees. The adult expansion group was the second largest, with about 25.5 million enrollees. Enrollment among COVID newly-eligible beneficiaries was substantially smaller than in the other major eligibility groups.

## Visualization 2: Enrollment Trends by Eligibility Group
```{r}
trend_data <- medicaid |>
  filter(
    Month >= as.Date("2017-01-01"),
    `majoreligibility group` != "Unknown"
  ) |>
  group_by(Month, `majoreligibility group`) |>
  summarize(
    enrollment = sum(CountEnrolled),
    .groups = "drop"
  )

trend_data |>
  ggplot(
    aes(
      x = Month,
      y = enrollment,
      color = `majoreligibility group`,
      group = `majoreligibility group`
    )
  ) +
  geom_line(linewidth = 0.9) +
  geom_point(
    data = trend_data |>
      filter(format(Month, "%m") == "12"),
    size = 2,
    alpha = 0.8
  ) +
  scale_x_date(
    breaks = as.Date(paste0(2017:2022, "-01-01")),
    date_labels = "%Y"
  ) +
  scale_y_continuous(
    labels = scales::label_number(
      scale = 1e-6,
      suffix = "M"
    )
  ) +
  labs(
    title = "Medicaid and CHIP Enrollment Trends Differ Across Eligibility Groups",
    subtitle = "Monthly enrollment across reporting jurisdictions, 2017–2022",
    x = NULL,
    y = "Enrollment",
    color = "Eligibility group",
    caption = "Source: Centers for Medicare & Medicaid Services (CMS)"
  ) +
  theme_minimal()
```

Children remained the largest Medicaid and CHIP eligibility group throughout 2017–2022. Enrollment increased notably after 2020 among children, adults, and especially the adult expansion group. In contrast, enrollment among aged beneficiaries and persons with disabilities changed more gradually over the period.

## Visualization 3

## Visualization 4

